\documentclass{article}

\begin{document}
\noindent\textbf{Name: Paul Gustafson}
\medskip

\noindent\textbf{Hidden subgroup problem for nilpotent groups of rank 2:} 


\begin{itemize}
\item Main paper:  http://arxiv.org/abs/0707.1260
\item HSP for extraspecial groups : https://arxiv.org/abs/quant-ph/0701235
\end{itemize}

\emph{Description.} We would like, at the very least, to reformulate the HSP algorithm in
of the main paper in a more accessible way.  Once we better understand how the algorithm works, 
we will evaluate whether we can extend it to groups of higher nilpotency class before the paper deadline.

\medskip

\noindent\textbf{Computing the Unit Group Algorithm for Number Fields:}

\begin{itemize}
\item Main paper:  http://dl.acm.org/citation.cfm?id=2591860
\end{itemize}

\emph{Description.} One of the main results of this paper is a solution to the
continuous HSP for $\mathbf{R}^n$, where the definition of the HSP is modified to assume that
the subgroup is a full-rank lattice and that the hiding function is Lipschitz,
periodic on the lattice, and only hides the lattice approximately.   

We would like to generalize their definition, and attempt to push their method of proof further.  
Lattices are also defined over locally compact topological groups.  This setting may be too broad for a good definition,
but we could restrict ourselves to, say, compact Lie groups, or even just try to apply it to a particular compact Lie group.
\medskip

\noindent\textbf{A Lattice Problem in Quantum NP:}


\begin{itemize}
\item Main paper:  https://arxiv.org/abs/quant-ph/0307220
\end{itemize}

\emph{Description.} We would like to analyze the importance of positive definite functions
in the context of the complexity class $QMA$.  We will summarize the results of the main paper,
and try to understand how if they can be used to generate solutions to lattice problems, not
just verify them.  To do this, we will consider the simplest possible lattice problems (in 
$\mathbf{R}^n$ for $n =1,2$).

\medskip

\noindent\textbf{Team partner:} Qing Zhang





 



\end{document}
