Reshitikhin - TQFT, semiclassical limits, integrable systems

Outline:
\begin{itemize}
\item Introduction
\item Representation theory and geometric quantization
\item q6j symbols - corresponding moduli spaces
\item Integrable systems
\item Semiclassical limits of integrable systems
\end{itemize}

Want to know asymptotics of 6j-symbols, related to semiclassical limits

A Liouville integrable systems on a symplectic manifold $(M, \omega)$ is 
a Lagrangian fibration.  Algebraically, corresponds to a maximal 
commutative Poisson subalgebra of $C(M)$.

Degenerate integrable system, composition of two
fibrations. Noncommutative, Poisson subalgebra.  Can look at center.
The center corresponds to Hamiltonians.  The rest corresponds to
conservation laws.

For the center, multiplicty of the joint spectrum grows as 
$$\hbar^{- \dim(fiber of \pi_2)/2}$$
as $\hbar \to 0$. 

Well-known example: hydrogen atom.

-------------

Deformation quantization

Formally,
$$f * g := f \cdot g +  \frac{ih}{2} \{f, g \} + O(h^2),$
where higher order terms are chosen so the product is associative.

-------------

Geometric quantization

$L$ - prequantum line bundle over $M$ with a connection $\alpha$, $\omega = d \alpha$

$F$ - polarization of $M$ (tangent distribution $F \subset (TM)_\CC$); 2 types real and holomorphic

Pick real polarization $F$

$$V = H^{(1/2)}_I (M)$$ covariantly constant $1/2$-densities

A density is a collection of functions on charts with a Jacobian multiplication on the chart overlaps

A $1/2$-density takes a square root of the Jacobian, so you can integrate the product of functions,
getting $L_2^{(1/2)}(M)$ for any manifold.


