\documentclass{article}
\usepackage{../m}

\begin{document}
\noindent Paul Gustafson\\
\noindent Texas A\&M University - Math 666\\ 
\noindent Instructor: Igor Zelenko

\subsection*{Bonus exercises}
\p{1} Let $\lambda:[0,T] \to [0,1]$ be measurable. Let 
$$A_n = \bigcup_{k=0}^{n-1} [kT/n, kT/n + \int_{kT/n}^{(k+1)T/n} \lambda(t) \,dt).$$
Show that for any $\phi \in L_1[0,T]$,
$$\int_0^t \chi_{A_n}(\tau) \phi(\tau) \,d\tau \to \int_0^t \lambda(\tau) \phi(\tau) \,d\tau,$$
uniformly on $[0,T]$.
\begin{proof}
We first consider the case $\phi = \chi_{(a,b)}$.  Pick nonnegative integers $k_a, k_b \le n$  such that $|k_aT/n - a| < T/n$ and $|b - k_bT/n| < T/n$.  Then 
\begin{align*}
\left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) \phi(\tau) \,d\tau \right| &= \left| \int_a^b (\lambda(\tau) - \chi_{A_n}(\tau)) \,d\tau \right|
\\ &= \left| \left( \int_a^{k_a T/n} + \int_{k_b T/n}^b + \int_{k_a T/n}^{k_b T/n} \right) (\lambda(\tau) - \chi_{A_n}(\tau)) \,d\tau \right|
\\ & \le |k_a T/n - a| (2) + |b - k_bT/n| (2) + 0
\\ & \le 4 T/n,
\end{align*}
which goes to $0$ uniformly in $t$.

By linearity, we get the same result for step functions. 

Let $\epsilon > 0$ and $\phi \in L_1[0,T]$ be arbitrary. We can pick a step function $h$ such that $\| \phi - h \|_{L_1[0,T]} < \epsilon/(2T)$.  Then
\begin{align*}
\left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) \phi(\tau) \, d\tau  \right| & \le \left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) h(\tau) \, d\tau  \right| + \left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) (h - \phi)(\tau) \, d\tau  \right|
\\ & \le \left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) h(\tau) \, d\tau  \right| + 2 t \epsilon/(2T)
\\ & \le \left| \int_0^t (\lambda(\tau) - \chi_{A_n}(\tau)) h(\tau) \, d\tau  \right| + \epsilon.
\\ & \le 2 \epsilon,
\end{align*}
uniformly in $t$ for $n$ sufficiently large by the step function case.
\end{proof}

\p{2} Let $Y_1$ and $Y_2$ be two complete vector fields on $\R^n$, which are also Lipshitzian, i.e. there exist $L_i > 0$ such that $\|Y_i(x_1) - Y_i(x_2)\| \le L_1 \|x_1 x_2 \|$ for any $x_1, x_2 \in \R^n$, $i \in \{1, 2\}$. Let $T, \lambda, A_n$ be as in the previous problem. Further, assume that $q(t)$ is the trajectory of the vector field $\lambda(t)Y_1 + (1 - \lambda(t)) Y_2$ with $q(0) = q_0$ and $(q_n(t))$ are the trajectories of the vector fields $\chi_{A_n} Y_1 + (1 - \chi_{A_n}(t))Y_2$ with $q_n(0) = q_n$ and such that $q_n \to q_0$. Prove that $q_n(t) \to q(t)$ uniformly on $[0,T]$.
(Hint: Use the Gronwall inequality: if $y(\tau), \beta(\tau)$ are continuous and take non-negative values on $[0,T]$ and $\alpha \ge 0$ with $y(t) \le \alpha + \int_0^t \beta(\tau) y(\tau) \, d\tau$ for $t \in [0,T]$, then $y(t) \le \alpha e^{\int_0^t \beta(\tau) \, d \tau}$.)


\begin{proof}
\end{proof}


\p{3} Let $V$ be a finite dimensional complex vector space and $T:V \to V$ be a linear transformation. Let $p(x)$ denote the characteristic polynomial of $T$, and $m(x)$ denote the minimal polynomial of $T$. Find a necessary and sufficient condition on the Jordan Normal Form of $T$ for $p(x) = m(x)$.
\begin{proof}
I claim that $p = m$ if and only if the geometric multiplicity of each eigenvalue of $T$ is $1$. This means that each Jordan block has a distinct eigenvalue.

Let $(\lambda_j)$ be an enumeration of the eigenvalues of $T$ without multiplicity. Let $s_j$ denote the size of the largest Jordan block corresponding to each eigenvalue $\lambda_j$. Then $(T - \lambda_j I)^{s_j}$ kills the generalized eigenspace $V_j$ for $\lambda_j$.   Thus $m(x)$ divides $\prod_j (x - \lambda_j)^{s_j}$.   Moreover, each $(x - \lambda_j)^{s_j}$ generates the $T$-annihilator for the basis vector acted on by the last column of the largest Jordan block for $\lambda_j$.   Thus, each $(x - \lambda_j)^{s_j}$ divides $m(x)$.  Hence $m(x) = \prod_j (x - \lambda_j)^{s_j}$.  It follows that $p = m$ iff each $\lambda_j$ corresponds to exactly one Jordan block.
\end{proof}
\end{document}
