\documentclass{article}
\usepackage{../m}

\begin{document}
\noindent Paul Gustafson\\
\noindent Math 689 Commutative and Homological Algebra

\subsection*{HW 1}
All rings are commutative with $1 \neq 0$.

\vskip 5pt

\p 1 Let $A$ be a ring and let $\mathfrak S$ be the set of all multiplicative subsets of $A$ that do not contain $0$, ordered by inclusion.  Prove that an element $S$ of $\mathfrak S$ is maximal in $\mathfrak S$ if and only if $R - S$ is a minimal prime ideal.
\begin{proof}
Suppose $S$ is a maximal element of $\mathfrak S$. By a proposition from class, there exists a prime ideal $I$ disjoint from $S$.  Suppose $J$ is a prime ideal contained in $I$.  Then $R - J$ is a multiplicative subset of $A$ not containing $0$, i.e, an element of $\mathfrak S$.  But we also have $S \subset R - I \subset R -J$. Thus, by the maximality of $S$, we have $S = R - I = R - J$.  Hence $I = J$. Since this for all prime ideals $J \subset I$, $I$ is a minimal prime ideal.

For the opposite implication, suppose $S$ is not a maximal element of $\mathfrak S$. Pick $T \in \mathfrak S$ with $S$ properly contained in $T$.  Pick a prime ideal $I$ disjoint from $T$. Then
$I \subset R - T \subset R - S$, where the last inclusion is proper.  Therefore, $R - S$ is not a minimal prime ideal.
\end{proof}

\p 2 Let $I$ and $J$ be ideals of a ring $A$. Show that 
$$\sqrt{IJ} = \sqrt{ I \cap J }   = \sqrt I \cap \sqrt J .$$

\begin{proof}
First we have $IJ \subset I \cap J$, so $\sqrt{IJ} \subset \sqrt{ I \cap J}$.

Similarly, $I \cap J \subset I$ and $I \cap J \subset J$ implies that $\sqrt{ I \cap J }   \subset \sqrt I \cap \sqrt J$.


Lastly, let $P$ be a prime ideal containing $IJ$.  I claim that $P$ contains $I$ or $J$. Suppose not.  Then there exists $x \in I - P$ and $y \in J - P$.
But then $xy \in IJ \subset P$, which contradicts the fact that $P$ is a prime ideal. Thus, $P$ contains $I$ or $J$. Thus every prime ideal containing $IJ$ contains either $I$ or $J$. Thus, $\sqrt I \cap \sqrt J \subset \sqrt{IJ}$.
\end{proof}

\p 3 Let $\phi: A \to B$ be a ring homomorphism, and $I$ an ideal of $B$. Prove that $\phi^{-1}(\sqrt I) = \sqrt{\phi^{-1}(I)}$.

\begin{proof}
Let $J \subset B$ be a prime ideal.  Let $a \in A$ and $x,y \in \phi^{-1}(J)$.  Then we have $\phi(x + y) = \phi(x) + \phi(y) \in J$, and $\phi(ax) = \phi(a)\phi(x) \in J$.  Thus $\phi^{-1}(J)$ is an ideal.   Now suppose $a,b \in A$ with $ab \in \phi^{-1}(J)$.  Then $\phi(a)\phi(b) = \phi(ab) \in J$,  hence $\phi(a)$ or $\phi(b)$ is in $J$ since $J$ is prime.  Thus, $\phi^{-1}(J)$ is also a prime ideal.  Since inverse maps preserve intersections and containment, this implies that $\phi^{-1}(\sqrt I) \supset \sqrt{\phi^{-1}(I)}$.

For the reverse inclusion, suppose $x^n \in \phi^{-1}(I)$.  Then $\phi(x^n) \in I$.  Hence $(\phi(x))^n \in I$, so $\phi(x) \in \sqrt I$.  Thus, $x \in \phi^{-1}(\sqrt I)$.
\end{proof}

\p 4 Let $n$ be a positive integer, $n \ge 2$. Find $\nil(\Z/(n))$.
\begin{proof}
The maximal ideals of $\Z/(n)$ are the ideals generated by the prime factors of $n$.  The intersection of these ideals, $\nil(\Z/(n))$,  is the ideal generated by the product of the distinct prime factors of $n$ taken without multiplicity.
\end{proof}

\p 5 Let $\phi: A \to B$ be a surjective ring homomorphism.
\begin{enumerate}[(a)]
\item Prove that $\phi(\rad(A)) \subset \rad(B)$.
\begin{proof}
Let $a \in \rad(A)$ and $b \in B$.  Since $\phi$ is surjective, there exists $c \in A$ with $\phi(c) = b$.  Since $a \in \rad(A)$, the element $1 - ca$ is a unit. Thus, $\phi(1 - ca) = 1 - b\phi(a)$ is a unit.  Since this holds for all $b \in B$, it follows that $\phi(a) \in \rad(B)$.  Since this holds for all $a \in \rad(A)$, we have $\phi(\rad(A)) \subset \rad(B)$.
\end{proof}
\item Give an example to show that the inclusion need not be an equality.
\begin{proof}
Let $\phi:\Z \to \Z/(4)$ be the canonical map.  We have $\rad(\Z) = (0)$ but $\rad(\Z/(4)) = (2)$.
\end{proof}
\end{enumerate}

\p 6 Let $A$ be a local ring.  Prove that $A$ contains no idempotent elements other than $0$ and $1$.
\begin{proof}
Suppose not.  Let $a \in A - \{0,1\}$ be idempotent.  Then $a(1-a) = 0$.  Thus $a$ and $1-a$ are both non-units. Thus $a$ lies in the unique maximal ideal of $A$, hence the Jacobson radical of $A$.  This contradicts the fact that $1-a$ is a non-unit.
\end{proof}

\p 7 Let $A$ be a local ring and $M,N$ finitely generated $A$-modules. Prove that if $M \otimes_A  N = 0$, then $M = 0$ or $N = 0$.
\begin{proof}
Let $k$ be the residue field of $A$, and $\mathfrak{m}$ the maximal ideal.  We have $0 = (M \otimes_A N) \otimes_A (k \otimes_A k) = (M \otimes_A k) \otimes_A (N \otimes_A k) = (M \otimes_A k) \otimes_k (N \otimes_A k)$.  The last expression is just a tensor product of finite dimensional vector spaces over the field $k$, so either $M \otimes_A k = 0$ or $N \otimes_A k = 0$.  WLOG suppose the former. Then $0 = M \otimes_A k = M/\mathfrak{m}M$, so by Nakayama's lemma $M = 0$.
\end{proof}

\p 8 Let $A$ be a ring. Suppose that $A^m \cong A^n$ as $A$-modules.  Show that $m = n$. (Hint: reduce to the case of a field.)
\begin{proof}
Let $\mathfrak{m}$ be a maximal ideal of $A$.  Then $A^m/\mathfrak{m}A^m \cong A^n/\mathfrak{m}A^n$ as $A/\mathfrak{m}A$-vector spaces. Since $A^m/\mathfrak{m}A^m \cong (A/\mathfrak{m}A)^m$ and the same for $n$, the invariance of dimension for vector spaces implies that $m = n$.
\end{proof}


\p 9 Let $I_1, \ldots, I_n$ be ideals of a ring for which $I_1 \cap \ldots \cap I_n = (0)$. Prove that if $A/I_j$ is Noetherian for each $j$, then $A$ is Noetherian.
\begin{proof}
Let $J \subset A$ be an ideal. Since $A/I_1$ is Noetherian, $J + I_1$ is a finitely generated ideal in $A/I_1$.  We can lift the generators to representatives in $J$ which generate some ideal $F_1 \subset J$.  We then have $J = F_1 + J \cap I_1$, where $F_1$ is finitely generated.

Similarly using $A/I_2$, we have $J \cap I_1 = F_2 + J \cap I_1 \cap I_2$, where $F_2 \subset J \cap I_2$  is finitely generated.  Continue in this way, we finally end up with $J \cap I_1 \cap \ldots \cap I_{n-1} = F_{n} + J \cap I_1 \cap \ldots \cap I_{n} = F_{n}$.

Substituting back, we get $J = F_1 + \ldots + F_{n}$, where the $F_i$ are all finitely generated.  Hence, $J$ is finitely generated.  Thus, $A$ is Noetherian.
\end{proof}

\p 10 Let $A, B, C$ be rings and suppose that $\phi: A \to C, \psi: B \to C$ are surjective ring homomorphisms. Prove that if $A$ and $B$ are Noetherian, then $A \times_C B$ is Noetherian. (Hint: use 9.).
\begin{proof}
Let $I,J \subset A \times_C B$ be the ideals defined by $I = A \times 0$ and $J = 0 \times B$. Then $I \cap J = 0$, and $A \times_C B/I = B$ and $A \times_C B/J = A$ which are Noetherian.  Hence, exercise 9 implies that $A \times_C B$ is Noetherian.
\end{proof}

\end{document}
