\documentclass{article}
\usepackage{../m}

\begin{document}
\noindent Paul Gustafson\\
\noindent Texas A\&M University - Math 641\\ 
\noindent Instructor - Fran Narcowich

\subsection*{HW 7}
\p{1} Let $f(x) = e^{x}$ for $-\pi < x < \pi$.

a. Find the complex form of the Fourier series for $f$.

b. Sketch three periods of the $2\pi$-periodic function to which the series converges pointwise. (Hand-drawn is fine. No need to use a computer here.)

c. Find $\sum_{n=0}^\infty \frac 1 {n^2 +1}$.

d. Estimate the error $\|f - S_N \|_{L_2[-\pi, \pi]}$, where $S_N$ is the partial sum of the Fourier series for $f$.

\begin{proof}
For (a), the Fourier series for $f$ is $\sum_{n = -\infty}^\infty a_n e^{inx}$ where
\begin{align*}
a_n &  = \frac 1 {2 \pi} \int_{-\pi}^{\pi} e^x e^{-inx} \, dx
\\ & =  \frac 1 {2 \pi (1 - in)} \left[ e^{(1 - in)x} \right]_{x = -\pi}^{\pi}
\\ & =  \frac {(-1)^n(e^{\pi} - e^{-\pi})} {2 \pi (1 - in)} 
\end{align*}

For (c), we have $|a_n|^2 = \frac{(e^{\pi} - e^{-\pi})^2}{(2 \pi)^2(1 + n^2)}$.  Hence using Parseval's identity, we have
\begin{align*}
\sum_{n=0}^\infty  \frac 1 {n^2 +1} & = \frac 1 2 + \frac 1 2 \sum_{n= -\infty}^\infty  \frac 1 {n^2 +1}
\\ & =\frac 1 2  + \frac{2 \pi^2}{(e^{\pi} - e^{-\pi})^2} \sum_{n = -\infty}^\infty |a_n|^2
\\ & =\frac 1 2  + \frac{ \pi}{(e^{\pi} - e^{-\pi})^2} \int_{-\pi}^{\pi} e^{2x} \, dx
\\ & =\frac 1 2  + \frac{ \pi}{(e^{\pi} - e^{-\pi})^2} \left(\frac 1 2 \right) (e^{2 \pi} - e^{-2\pi})
\\ & =\frac 1 2  + \frac{ \pi (e^{ \pi} + e^{-\pi}) }{2(e^{\pi} - e^{-\pi})}  
\\ & = \frac 1 2 + \frac \pi 2 \coth(\pi)
\end{align*}

For (d), we have 
\begin{align*}
\|f - S_N \|_{L_2[-\pi, \pi]}^2 & = \|\sum_{|n| > N} a_n e^{inx} \|_{L_2[-\pi, \pi]}^2
\\ & = \sum_{|n| > N} |a_n|^2
\\ & =  \frac{(e^{\pi} - e^{-\pi})^2}{(2 \pi)^2} \sum_{|n| > N}(1 + n^2)^{-1}
\\ & =  \frac{(e^{\pi} - e^{-\pi})^2}{(2 \pi)^2} \sum_{|n| > N} \int_{n-1}^n (1 + t^2)^{-1} \, dt
\\ & =  \frac{(e^{\pi} - e^{-\pi})^2}{2 \pi^2 } \int_{N}^\infty \frac{dt}{1 + t^2}
\\ & =  \frac{(e^{\pi} - e^{-\pi})^2}{2 \pi^2 } \left( \frac \pi 2  - \tan^{-1}(N)  \right)
\end{align*}
\end{proof}


\p{2} Prove this:  Let $g$ be a $2 \pi$-periodic piecewise continuous function. Then, $\int_{-\pi+c}^{\pi + c} g(u) \, du$ is independent of $c$. (Remark: This holds for $g$ integrable on each bounded interval of $\R$.)

\begin{proof}
Pick $k \in \Z$ such that $2\pi k \in [-\pi + c, \pi + c)$.  We have 
\begin{align*}
\int_{-\pi+c}^{\pi + c} g(u) \, du & = \int_{-\pi + c}^{2\pi k} g(u) \, du +  \int_{2\pi k}^{\pi + c} g(u) \, du
\\ & = \int_{-2\pi (k-1) - \pi + c}^{2\pi} g(v + 2 \pi (k-1))  \, dv + \int_0^{\pi + c - 2 \pi k} g(v + 2 \pi k) \, dv
\\ & = \int_{-2\pi k + \pi + c}^{2\pi} g(v)  \, dv + \int_0^{\pi + c - 2 \pi k} g(v) \, dv
\\ &  = \int_0^{2\pi} g(v) \, dv.
\end{align*}
\end{proof}


\p{3} Use the previous result to show that if $f$ is $2 \pi$-periodic and piecewise smooth, then it has the Fourier series $f(x) \sim a_0 + \sum_{n=1}^\infty a_n \cos(nx) + b_n \sin(nx)$ where
$$ a_0 = \frac 1 {2\pi} \int_0^{2\pi} f(x) \, dx \quad a_n = \frac 1 \pi \int_0^{2\pi} f(x) \cos(nx) \, dx \quad b_n = \frac 1 \pi \int_0^{2\pi} f(x) \sin(nx) \, dx. $$
Formulate a theorem on the pointwise convergence of the series.

\begin{proof}
The first part is just a rewriting of the definition of Fourier series.  The theorem about pointwise convergence is that the series converges pointwise to $f$. This was proved in class by applying the Riemann-Lebesgue lemma to the convolution of $f$ with the Dirichlet kernel.
\end{proof}


\p{4} Find the Fourier series for $f(x)=x$, $0<x<2 \pi$. Sketch three periods of the $2\pi$-periodic function to which the series converges pointwise. (Hand-drawn is fine. No need to use a computer here.)

\begin{proof}
The Fourier series for $f$ is $\sum_{n = -\infty}^{\infty} a_n e^{inx}$ where $a_0 = \frac 1 {2\pi} \int_{0}^{2 \pi} x \,dx = \frac 1 2$, and for $n \neq 0$
\begin{align*}
a_n & = \frac 1 {2 \pi} \int_0^{2\pi} x e^{-inx} \, dx
\\ & = \frac 1 {2 \pi} \left[ x \left( \frac{-1}{in} \right) e^{-inx} \right]_0^{2\pi} + \frac 1 {2\pi in} \int_0^{2\pi} e^{-inx} \, dx
\\ & = \frac 1 {2 \pi} \left( 2\pi \left( \frac{-1}{in} \right)  \right) + \frac 1 {2\pi in} \left[ \frac {-1}{in} e^{-inx} \right]_0^{2\pi}
\\ & = \frac i n.
\end{align*}
\end{proof}

\p{5} Find the Fourier series for $\displaystyle f(x)=\left\{ \begin{array}{ll}
1, & x \in [- \frac \pi 4 , \frac \pi 4]
\\ 0, & x \in (-\pi,- \frac \pi 4) \cup(\frac \pi 4, \pi) 
\end{array}\right. .$
\begin{proof}
The Fourier series is $\sum_{n = -\infty}^{\infty} a_n e^{inx}$ where
\begin{align*}
a_n & = \frac 1 {2 \pi} \int_{-\pi/4}^{\pi / 4} e^{-inx} \, dx
\\ & =  \frac 1 {2 \pi} \left[ - \frac {e^{-inx}}{in} \right]_{-\pi/4}^{\pi/4}
\\ & =   \frac 1 {\pi n} \sin(n \pi/4)
\end{align*}

\end{proof}

\p{6} Consider the series $\sum_{n=-\infty}^\infty c_n e^{inx}$, where $\sum_{n=-\infty}^\infty |c_n| < \infty$. Show that the series $\sum_{n= -\infty}^\infty c_n e^{inx}$ converges uniformly to a $2\pi$-periodic continuous function $f(x)$ and the series is the Fourier series for $f$. Also, show that the series converges to $f$ in $L_2[-\pi, \pi]$.

\begin{proof}
For $x \in \R$, we have $\sum_{n = -\infty}^\infty |c_n e^{inx}| = \sum_n |c_n| < \infty$. Hence $\sum_{n = -\infty}^\infty c_n e^{inx}$ converges absolutely for all $x$ and uniformly in $x$.  Since $f$ is the uniform limit of continuous functions, it is continuous. It is obvious that $f$ is  $2\pi$-periodic and that the series is the Fourier series for $f$. 

Let $f_N = \sum_{n = -N}^N c_n e^{inx}$.  Then $\int_{-\pi}^\pi |f_N - f|^2 \,dx \le 2 \pi \|f_N - f\|_\infty^2 \to 0$. Hence $f_N \to f$ in $L_2[-\pi, \pi]$.
\end{proof}

\end{document}
